---
title: "Exercise set 1"
author: "Claudi Lleyda Moltó"
output:
  pdf_document: default
editor_options:
  chunk_output_type: console
---
```{r}
# Claudi Lleyda Moltó
# NIU: 1462908
```

```{r}
#setwd("~/UAB/MSE/Data-visualisation/Exercises/intro-to-R/ex1/")
#ls()
```

Exercise 1
==========

(a) Enter the data into R, as two vectors

```{r}
year <- 2012:2021
snow.cover <- c(47.5, 49.2, 46.0, 47.3, 48.8, 49.2, 46.9, 47.2, 46.4, 46.8)
```

(b) Take the logarithm of `snow.cover` and call it `log.snow.cover`

```{r}
log.snow.cover <- log(snow.cover)
```

(c) Plot `snow.cover` versus `year`

```{r}
plot(year, snow.cover, "l")
```

(d) Create a summary of the snow cover data

```{r}
summary(snow.cover)
```

Exercise 2
==========

```{r}
set.seed(1462908)

x <- sample(1:6, 100, replace = TRUE)
```

(a) Count the number of 3’s in `x`

```{r}
sum(x == 3)
```

(b) What is the value of the 55th component of `x`?

```{r}
x[55]
```

(c) Display the components of `x` corresponding to odd positions from 11th to 19th

```{r}
x[seq(11, 19, by = 2)]
```

(d) Modify `x` so that 3’s are changed to 1

```{r}
x[x == 3] <- 1
```

(e) Create a vector `y` with the components of `x` that are greater than 2

```{r}
y <- x[x > 2]
````

(f) Calculate the mean of the square root of `x`

```{r}
mean(sqrt(x))
```

(g) Concatenate `x` and `y` in a new vector called `z`

```{r}
z <- c(x, y)
```

(h) Remove `z` from your workplace.

```{r}
rm(z)
```

Exercise 3
==========

```{r}
subletters <- letters[1:5]
subnums <- 1:3

paste(rep(subletters, each = 3), subnums, sep = "")
paste(subletters, rep(subnums, each = 5), sep = "")
```

Exercise 4
==========

```{r}
rincome <- scan("data/rincome.txt", character(), sep="\n", skipNul = T)
```

(a) Convert the values `"Don't know"` `"No answer"` `"Not applicable"` `"Refused"` into `NA`

```{r}
toNA = c("Don't know", "No answer", "Not applicable", "Refused")
rincome[rincome %in% toNA] <- NA
```

(b) Create a factor `rincome.f` with the values of `rincome` into a factor and display its levels

```{r}
rincome.f <- factor(rincome, ordered = FALSE)
levels(rincome.f)
```

(c) Create a new vector `rincome.lv` with the levels of `rincome.f`

```{r}
rincome.lv <- levels(rincome.f)
```

(d) Reorder `rincome.lv` so that `"Lt $1000"` is the first

```{r}
rincome.lv <- levels(relevel(factor(rincome.lv), "Lt $1000"))
rincome.lv
```

(e) Create an ordered factor `rincome.o` with `rincome.f` and the order of levels given by `rincome.lv`

```{r}
rincome.o <- ordered(factor(rincome, levels = rincome.lv))
```

(f) With `summary(rincome.o)` see the result of the previous manipulation

```{r}
summary(rincome.o)
```

(g) Apply the function class to all the objects in this exercice to see the differences

```{r}
class(rincome)
class(rincome.f)
class(rincome.lv)
class(rincome.o)
```
